Testing SEACells results

### 2024-11-07 analyses/metacells/notebooks/testing-seacells.Rmd
update testing notebook and add rendering to script

---
title: "Testing SEACells results"
output: html_notebook
---

@@

```{python imports}
import pathlib

@@

```{r paths}
root_dir <- rprojroot::find_root(rprojroot::is_git_root)
data_dir <- file.path(root_dir, "data", "current")

infile <- file.path(data_dir, "SCPCP000001/SCPCS000001/SCPCL000001_processed_rna.h5ad")
```

## Load data
```{python load_data}
ad = sc.read(r.infile)

# put the highly variable genes in a column of adata.var
ad.var["highly_variable"] = ad.var.gene_ids.isin(ad.uns["highly_variable_genes"])
```

## Run SEACells

```{python run_seacells}
n_SEACells = round(ad.n_obs / 75)
n_waypoint_eigs = 10

model = SEACells.core.SEACells(
    ad,
    build_kernel_on="X_pca",
    n_SEACells=n_SEACells,
    n_waypoint_eigs=n_waypoint_eigs,
    convergence_epsilon=1e-5,
)
```
```{python}
model.construct_kernel_matrix()
M = model.kernel_matrix
```

```{python}
sns.clustermap(M.toarray()[:500, :500])
plt.show()
```

```{python}
model.initialize_archetypes()
SEACells.plot.plot_initialization(ad, model)
```

```{python}
model.fit(min_iter=10, max_iter=50)
model.plot_convergence()
```

```{python}
SEACells.plot.plot_2D(ad, key="X_umap", colour_metacells=True)
```

```{python}
SEACells.plot.plot_SEACell_sizes(ad, bins=10)
```

```{python}
purity = SEACells.evaluate.compute_celltype_purity(
    ad, "singler_celltype_ontology"
)
```

>>>>

---
title: "Testing SEACells results"
output: html_notebook
params:
  project_id: "SCPCP000001"
  sample_id: "SCPCS000001"
  library_id: "SCPCL000001"
---

@@

```{python imports}
import pathlib
import pickle

@@

```{r paths}
root_dir <- rprojroot::find_root(rprojroot::is_git_root)
data_dir <- file.path(root_dir, "data", "current")
base_dir <- file.path(root_dir, "analyses", "metacells")
results_dir <- file.path(base_dir, "results", params$project_id, params$sample_id)

ad_file <- file.path(data_dir, params$project_id, params$sample_id, paste0(params$library_id, "_processed_rna.h5ad"))
sc_anndata_file <- file.path(results_dir, paste0(params$library_id, "_seacells.h5ad"))
sc_model_file <- file.path(results_dir, paste0(params$library_id, "_seacells_model.pkl"))

```

## Load SEACells results

```{python load_data}
ad = sc.read(r.ad_file)
sc_ad = sc.read(r.sc_anndata_file)
with open(r.sc_model_file, "rb") as f:
  sc_model = pickle.load(f)
```

```{python}
_ = sns.clustermap(sc_model.kernel_matrix.toarray()[:500, :500])
plt.show()
```

```{python}
SEACells.plot.plot_initialization(sc_ad, sc_model)
```

```{python}
sc_model.plot_convergence()
```

```{python}
SEACells.plot.plot_2D(sc_ad, key="X_umap", colour_metacells=True)
```

```{python}
SEACells.plot.plot_SEACell_sizes(sc_ad, bins=10)
```

```{python}
purity = SEACells.evaluate.compute_celltype_purity(
    sc_ad, "singler_celltype_ontology"
)
```
